### 2020-04-18 serial.tex
tweaks

The generation interval is defined as the time between when an individual (infector) is infected and when an individual infects another person (infectee);
the generation-interval distributions plays a key role in shaping the relationship between the exponential growth rate $r$ and the reproduction number $\mathcal R$ \citep{wallinga2007generation}.
On the other hand, the serial interval is defined as the time between when an infector and an infectee become symptomatic \citep{svensson2007note}.
While generation and serial intervals are conceptually similar,
previously studies have often alluded to differences in their variances, despite having seemingly identical expected values \citep{svensson2007note,klinkenberg2011correlation,champredon2018equivalence};
some suggested that using serial intervals can give different estimates of $\mathcal R$ \citep{britton2019estimation}.
Even though these distinctions were first made over a decade ago \citep{svensson2007note}, 
the need for a better conceptual and theoretical framework for understanding their differences is becoming clearer as the COVID-19 pandemic unfolds:
Researchers continue to rely on both generation and serial intervals to estimate $\mathcal R$ for COVID-19 without making a clear distinction.

The lack of clear understanding of their differences can be largely attributed to an apparent paradox.
When the epidemic is growing exponentially, the spread of infection can be characterized as a ``renewal process'' based on previous incidence of infection, the associated generation-interval distribution, and the average infectiousness of an infected individual.
This renewal formulation allows us to link the exponential growth rate of an epidemic $r$ with its reproduction number $\mathcal R$.
Likewise, since infection typically leads to symptoms, we should be able to describe the renewal process of symptomatic cases using the serial-interval distribution.
Therefore, both generation- and serial-interval distributions should give us identical estimates of  $\mathcal R$ based the observed epidemic growth rate $r$.
However, current theory suggests that using generation- and serial-interval distributions give different estimates of $\mathcal R$ and contradicts our biological intution.

Here, we provide an answer to the decade-old paradox between generation and serial intervals.
We present a novel framework for understanding serial intervals and show that the serial-interval distribution must be defined with reference to disease incidence;
previous frameworks implicitly assume that incidence remains constant.
We are show that both serial- and generation-interval distributions can give identical estimates of $\mathcal R$ based on the exponential growth phase of an epidemic if they are properly defined.
However, failing to account for changes in the observed serial-interval distributions over time can severely bias the estimate of $\mathcal R$.
We apply our framework to serial intervals of COVID-19 and lay out several principles to consider in using information about serial intervals and other epidemiological time delays in the analysis of the ongoing pandemic.

@@

We first begin by describing a general framework for characterizing a distribution of time delays between \emph{any} two epidemiological events;
these events can be defined either within an infected individual (e.g., infection and symptom onset of an individual) or between infected individuals (e.g., symptom onsets of an infector and an infectee).
We can further divide these events into \emph{primary} and \emph{secondary} events.
When we measure an epidemiological time delay within an infected individual (e.g., incubation period), the primary event always occurs before the secondary event.
When we measure an epidemiological time delay between infected individuals (e.g., serial interval), 
the primary and secondary events are defined in terms of the direction of transmission:
The primary event refers to the event that occurs within an infector and does not necessarily occur before the secondary event.

We model time delays between a primary and a secondary event from a cohort perspective.
A primary cohort consists of \emph{all} individuals whose primary event occurred at a given time; 
a secondary cohort can be defined similarly based on the secondary events.
For example, when we are measurign serial intervals, a primary cohort $s$ consists of all infectors who became symptomatic at time $s$.
Then, for each primary cohort $s$, we can define the expected time distribution between primary and secondary events.
We refer to this distribution as the forward delay distribution and denote it as $f_s(\tau)$.
The forward delay distributions can vary across primary cohorts.

Likewise, we can define the backward delay distribution $b_s(\tau)$ for a secondary cohort $s$:
The backward delay distribution for secondary host $s$ describes the time delays between a primary and secondary host given that the secondary event occurred at time $s$.
It follows that:
\begin{equation}
b_s(\tau) \propto C(s-\tau) f_{s-\tau}(\tau)
\end{equation}
where $C(s)$ is the size of the primary cohort $s$.
Therefore, changes in the backward delay distribution depends on the changes in cohort size $C(s)$ (therefore incidence of infection) as well as changes in the forward delay distribution.
These ideas apply to all epidemiological delay distributions and generalize the work by \citep{champredon2015intrinsic} who compared forward and backward generation-interval distributions to describe the realized generation intervals from the perspective of an infector and an infectee, respectively.

@@

The serial interval is defined as the time between when an infector becomes symptomatic and when and infectee becomes symptomatic.
Serial intervals $\tau$ are typically written as:
\begin{equation}
\tau = - x_0 + \sigma + x_1
\end{equation}
where $x_0$ and $x_1$ represent the realized time from infection to symptom onset of an infector and an infectee, respectively, and $\sigma$ represents the realized generation interval.
Previous studies have often assumed that $x_0$ and $x_1$ follow the same distributions and concluded that the serial and generation intervals have the same mean \citep{svensson2007note,klinkenberg2011correlation,champredon2018equivalence, britton2019estimation};
however, they implicitly assume that incidence stays constant.

@@

Using the cohort-based framework provides a clear way of understanding the serial-interval distribution.
Given that an infector became symptomatic at time $\ell$, we have to first go backward in time by asking when the infector was infected and go forward in time by asking when the infector infected infectee and when the infectee became symptomatic;
this defines the forward serial interval (\fref{diagram}A).
Then, it is clear that $x_0$ represents the backward incubation period of infectors who became symptomatic at time $\ell$; 
$\sigma$ represents the forward generation interval for individuals who came infected at time $\ell - x_0$ conditional on their incubation period $x_0$;
and $x_1$ represents the forward incubation period for the infectees who became infected at time $\ell - x_0 + \sigma$.
Assuming that the forward incubation distribution does not vary across cohorts, the forward serial interval distribution of a primary cohort $\ell$ can be written as follows:
\begin{equation}
f_\ell(\tau) \propto \int_{0}^\infty \int_{0}^\infty \mathcal R_c (\ell - x_0) i(\ell - x_0) h_{\ell - x_0}(x_0, \sigma) k(\tau-\sigma+x_0) \mathrm{d} x_0\, \mathrm{d}\sigma
\label{eq:forward}
\end{equation}
where $i$ is the incidence of infection, $h$ is the joint probability distribution of the forward incubation period and forward generation interval of a primary cohort $\ell - x_0$, and $k$ is a marginal probability distribution of $h$ describing the forward incubation periods:
\begin{equation}
k(x_0) = \int_0^\infty h_\ell(x_0, \sigma) \mathrm{d}\sigma.
\end{equation}
Finally, $\mathcal R_c (\ell - x_0)$ is the cohort reproduction number, which is defined as the average secondary number of cases caused by an individual infected at time $\ell - x_0$.
We have to multiply the joint forward distribution $h$ by $\mathcal R_c (t) i(t)$, which represents total infectiousness of individuals infected at time $t$, rather than incidence $i(t)$, because we are interested in the backward incubation period $x_0$ and the forward generation interval $\sigma$ of infectors (i.e., those who transmitted successfully to at least one individual), rather than all infected individuals.
We revisit this idea in the Results section.

@@

Here, use a renewal equation to model the spread of disease in a population.
Ignoring the dynamics of births and deaths, changes in the proportion of susceptible individuals $S(t)$ and incidence of infection $i(t)$ can be written as:
\begin{equation}
\begin{aligned}
\frac{\mathrm{d}S}{\mathrm{d}t} &= - i(t)\\
i(t) &= \mathcal R_0 S(t) \int_0^\infty i(t-\tau) g(\tau) \mathrm{d}\tau,
\end{aligned}
\label{eq:renewal}
\end{equation}
where $\mathcal R_0$ is the basic reproduction number, and $g(\tau)$ is the intrinsic generation-interval distribution (i.e., the forward generation-interval distribution of a primary case in a fully susceptible population; \cite{champredon2015intrinsic}).
Then, the forward generation-interval for a primary cohort $\ell$ follows \citep{champredon2015intrinsic}:
\begin{equation}
g_\ell (\tau) \propto g(\tau) S(\ell + \tau),
\end{equation}
which allows us to separate the joint probability distribution $h_\ell$ of the forward incubation period and the forward generation-interval distribution as a product of joint probability distribution $h$ of the forward incubation period and the intrinsic generation intervals and the proportion of susceptible individuals $S$:
\begin{equation}
h_\ell (x_0, \tau) \propto h(x_0, \tau) S(\ell + \tau),
\end{equation}
which satisfies the following:
\begin{equation}
g(\tau) = \int_0^\infty h(x_0, \tau) \mathrm{d}x_0.
\end{equation}
Finally, the cohort reproduction is defined as follows:
\begin{equation}
\mathcal R_c(t) = \mathcal R_0 \int_0^\infty g(\tau) S(t+\tau) \mathrm{d} \tau.
\end{equation}

@@

First, we compare the estimates of the reproduction number $\mathcal R$ based on the intrinsic generation-interval distribution $g(\tau)$ and the forward serial-interval distribution during the exponential growth phase $f_{\textrm{\tiny exp}}(\tau)$ (\fref{rR}A).
Both distributions provide identical estimates of $\mathcal R$ regardless of the correlation $\rho$ between the incubation period distribution and the intrinsic generation-interval distribution.
On the other hand, using naive serial-interval distributions that do not account for disease dynamics (i.e., assuming that the backward and the forward incubation period distributions are identical) underestimates $\mathcal R$;
as $r$ increases, ${\mathcal R}_{\textrm{\tiny naive}}$ saturates and eventually decreases due to negative serial intervals (\fref{rR}B).
While the forward serial intervals during the exponential growth phase can be also negative, the proportion of negative intervals are appropriately balanced because faster epidemic growth will lead to shorter backward incubation period (and therefore less negative serial intervals).

@@

\begin{figure}[!th]
\includegraphics[width=\textwidth]{forward.pdf}
\caption{
\textbf{Epidemiological dynamics and changes in mean forward and backward delay distributions.}
A. Cumulative incidence over time.
B-D. Changes in the mean forward incubation period, generation interval, and serial interval.
E-G. Changes in the mean backward incubation period, generation interval, and serial interval.
Black lines represent the results of a deterministic simulation.
Orange lines and points represent the results of a stochastic simulation.
Forward incubation periods and intrinsic generation-intervals are assumed to be independent of each other: $h(x_0, \sigma) = k(x_0) g(\sigma)$.
See Table 1 for parameter values.
}
\label{fig:epi}
\end{figure}

\fref{epi} compares the epidemiological dynamics (A) with the mean forward (B--D) and the mean backward (E--F) delay distributions of a deterministic model based on the renewal equation (\eref{renewal}) and the corresponding stochastic realization based on the Gillespie algorithm.
The mean forward incubation period remains constant throughout an epidemic as expected (\fref{epi}B).
The mean forward generation interval contracts as the epidemic progresses because an infected individual is less likely to infect another person as the proportion of susceptible individuals decreases (\fref{epi}C; \cite{champredon2015intrinsic}).
In contrast, the mean serial interval depends on previous incidence and therefore decreases over time (\fref{epi}D):
When incidence is increasing, symptomatic individuals are more likely to have been infected more recently (shorter backward incubation period and therefore longer forward serial interval) whereas when incidence is decreasing, symptomatic individuals are more likely to have been infected later (longer backward incubation period and therefore shorter forward serial interval).
Qualitative patterns in the changes in the mean backward delays is robust across all delay distributions because they are predominantly driven by the changes in incidence (\fref{epi}D).

\begin{figure}[!th]
\includegraphics[width=\textwidth]{forward_tease.pdf}
\caption{
\textbf{Dynamical correlations between incubation periods and the number of secondary cases.}
A. The mean forward serial interval depends on changes in the cohort reproduction number.
B. The mean backward incubation period of infectors is higher than the mean backward incubation period of all symptomatic cases.
C. Correlation between the number of secondary cases and the incubation period for individuals who got infected within the same time preiod.
D. Correlation between the number of secondary cases and the incubation period for individuals who developed symptoms within the same time preiod.
}
\label{fig:tease}
\end{figure}

We further tease apart factors that affect the forward serial interval distribution.
\fref{tease}A compares the mean forward serial interval calculated from a stochastic model (orange points) with their deterministically calculated values (black lines);
failing to account for changes in $\mathcal R_c$ in \eref{forward} overestimates the mean forward serial interval.
A key distinction that needs to be made is that the forward serial-interval distributions depends on the backward incubation period of symptomatic infectors (i.e., those who were able to successfully spread the disease to another person), rather than the incubation period of all symptomatic cases (\fref{tease}B).
Therefore, the cohort size of infectors must be calculated as a product of the cohort reproduction number and incidence.

>>>>

The generation interval is defined as the time between when an individual (infector) is infected and when an individual infects another person (infectee);
the generation-interval distributions plays a key role in shaping the relationship between the exponential growth rate $r$ and the reproduction number $\mathcal R$ \citep{wallinga2007generation}.
Similarly, the serial interval is defined as the time between when an infector and an infectee become \emph{symptomatic} \citep{svensson2007note}.
While serial intervals are similar to generation intervals, previous studies have noted that, in many contexts, serial intervals are expected to have larger variances than generation intervals but have the same mean \citep{svensson2007note,klinkenberg2011correlation,champredon2018equivalence};
some studies further suggested that using serial intervals can give different estimates of $\mathcal R$ \citep{britton2019estimation}.
Even these distributions were clearly distinguished over a decade ago \citep{svensson2007note}, 
the need for a better conceptual and theoretical framework for understanding their differences is becoming clearer as the COVID-19 pandemic unfolds:
Researchers continue to rely on both generation and serial intervals to estimate $\mathcal R$ for COVID-19 without making a clear distinction.

One important source of confusion comes down to an apparent paradox.
When the epidemic is growing exponentially, the spread of infection can be characterized as a ``renewal process'' based on previous incidence of infection, the associated generation-interval distribution, and the average infectiousness of an infected individual.
This renewal formulation allows us to link the exponential growth rate of an epidemic $r$ with its reproduction number $\mathcal R$ \citep{wallinga2007generation}.
Likewise, we should be able to describe the renewal process of symptomatic cases using the serial-interval distribution.
Therefore, both generation- and serial-interval distributions should give us identical estimates of  $\mathcal R$ based the observed epidemic growth rate $r$.
In contexts where the distributions are expected to be different, current theory has no explanation for how they might link $r$ to \Rx\ in the same way.

Here, we provide an answer to this paradox, by showing that the relevant interval for the renewal framework is what is called the ``forward'' interval, and that the forward serial (but not generation) interval is affected by the rate of growth $r$ during the exponential-growth phase.
We develop a new framework for understanding serial intervals and show that the forward serial-interval distribution gives the correct value of $\mathcal R$.
\jd{Come back to this. Not just over time, right?}
However, failing to account for changes in the observed serial-interval distributions over time can severely bias the estimate of $\mathcal R$.
We apply our framework to serial intervals of COVID-19 and lay out several principles to consider in using information about serial intervals and other epidemiological time delays in the analysis of the ongoing pandemic.

@@

We first begin by describing a general framework for characterizing a distribution of time delays between two epidemiological events;
these events can be defined either within an infected individual (e.g., infection and symptom onset of an individual, the incubation period) or between infected individuals (e.g., symptom onsets of an infector and an infectee, the serial interval).
We can further divide these events into \emph{primary} and \emph{secondary} events.
When we measure an epidemiological time delay within an infected individual (e.g., the incubation period), the primary event always occurs before the secondary event.
\jd{Rethink, or else rethink ``any'' above: the ``hidden'' time (my term) is the time between onset of infectiousness and onset of symptoms; it is within, but can be positive or negative. I'm not saying we should use it here\ldots}
When we measure an epidemiological time delay between infected individuals (e.g., the serial interval), 
the primary and secondary events are defined in terms of the direction of transmission:
The primary event refers to the event that occurs within an infector and does not necessarily occur before the secondary event.

We model time delays between a primary and a secondary event from a cohort perspective.
A primary cohort consists of \emph{all} individuals whose primary event occurred at a given time; 
a secondary cohort can be defined similarly based on the secondary events.
For example, when we are measuring serial intervals, a primary cohort $s$ consists of all infectors who became symptomatic at time $\ell$.
Then, for each primary cohort $\ell$, we can define the expected time distribution between primary and secondary events.
We refer to this distribution as the forward delay distribution and denote it as $f_\ell(\tau)$.
The forward delay distributions can vary across primary cohorts.

Likewise, we can define the backward delay distribution $b_\ell(\tau)$ for a secondary cohort $\ell$:
The backward delay distribution for secondary host $\ell$ describes the time delays between a primary and secondary host given that the secondary event occurred at time $\ell$.
It follows that:
\begin{equation}
b_\ell(\tau) \propto C(\ell-\tau) f_{\ell-\tau}(\tau)
\end{equation}
where $C(\ell)$ is the size of the primary cohort $\ell$.
Therefore, changes in the backward delay distribution depends on the changes in cohort size $C(\ell)$ (therefore incidence of infection) as well as changes in the forward delay distribution.
These ideas apply to all epidemiological delay distributions and generalize the work by \citep{champredon2015intrinsic} who compared forward and backward generation-interval distributions to describe the realized generation intervals from the perspective of an infector and an infectee, respectively.

@@

The serial interval is defined as the time between when an infector becomes symptomatic and when and infectee becomes symptomatic.
Serial intervals $\tau$ are typically written as:
\begin{equation}
\tau = - x_0 + \sigma + x_1
\end{equation}
where $x_0$ and $x_1$ represent the realized time from infection to symptom onset of an infector and an infectee, respectively, and $\sigma$ represents the realized generation interval.
Previous studies have often assumed that $x_0$ and $x_1$ follow the same distributions and concluded that the serial and generation intervals have the same mean \citep{svensson2007note,klinkenberg2011correlation,champredon2018equivalence, britton2019estimation};
however, these models implicitly assume that incidence stays constant.

@@

Using the cohort-based framework provides a clear way of understanding the serial-interval distribution.
Given that an infector became symptomatic at time $\ell$, we have to first go backward in time by asking when the infector was infected and then go forward in time by asking when the infector infected infectee and when the infectee became symptomatic;
this defines the forward serial interval (\fref{diagram}A).
Then, it is clear that $x_0$ represents the backward incubation period of infectors who became symptomatic at time $\ell$; 
$\sigma$ represents the forward generation interval for individuals who came infected at time $\ell - x_0$ conditional on their incubation period $x_0$;
and $x_1$ represents the forward incubation period for the infectees who became infected at time $\ell - x_0 + \sigma$.
Assuming that the forward incubation distribution does not vary across cohorts, the forward serial interval distribution of a primary cohort $\ell$ can be written as follows:
\begin{equation}
f_\ell(\tau) \propto \int_{0}^\infty \int_{0}^\infty \mathcal R_c (\ell - x_0) i(\ell - x_0) h_{\ell - x_0}(x_0, \sigma) k(\tau-\sigma+x_0) \mathrm{d} x_0\, \mathrm{d}\sigma
\label{eq:forward}
\end{equation}
where $i$ is the incidence of infection, $h$ is the joint probability distribution of the forward incubation period and forward generation interval of a primary cohort $\ell - x_0$, and $k$ is a marginal probability distribution of $h$ describing the forward incubation periods:
\begin{equation}
k(x_0) = \int_0^\infty h_\ell(x_0, \sigma) \mathrm{d}\sigma.
\end{equation}
Finally, $\mathcal R_c (\ell - x_0)$ is the cohort reproduction number, which is defined as the average secondary number of cases caused by an individual infected at time $\ell - x_0$.
We have to multiply the joint forward distribution $h$ by $\mathcal R_c (t) i(t)$, which represents total infectiousness of individuals infected at time $t$, rather than incidence $i(t)$, because we are interested in the backward incubation period $x_0$ and the forward generation interval $\sigma$ of infectors (i.e., those who transmitted successfully to at least one individual), rather than all infected individuals.
We revisit this idea in the Results section.

@@

Here, we use a renewal equation to model the spread of disease in a population \citep{heesterbeek1996concept, diekmann2000mathematical, roberts2004modelling, aldis2005integral, roberts2007model, champredon2018equivalence}.
Ignoring the dynamics of births and deaths, changes in the proportion of susceptible individuals $S(t)$ and incidence of infection $i(t)$ can be written as:
\begin{equation}
\begin{aligned}
\frac{\mathrm{d}S}{\mathrm{d}t} &= - i(t)\\
i(t) &= \mathcal R_0 S(t) \int_0^\infty i(t-\tau) g(\tau) \mathrm{d}\tau,
\end{aligned}
\label{eq:renewal}
\end{equation}
where $\mathcal R_0$ is the basic reproduction number, and $g(\tau)$ is the intrinsic generation-interval distribution (i.e., the forward generation-interval distribution of a primary case in a fully susceptible population; \cite{champredon2015intrinsic}).
Then, the forward generation-interval for a primary cohort $\ell$ follows \citep{champredon2015intrinsic}:
\begin{equation}
g_\ell (\tau) \propto g(\tau) S(\ell + \tau),
\end{equation}
which allows us to separate the joint probability distribution $h_\ell$ of the forward incubation period and the forward generation-interval distribution as a product of the proportion of susceptible individuals $S$ and the joint probability distribution $h$ of the forward incubation period and the intrinsic generation intervals:
\begin{equation}
h_\ell (x_0, \tau) \propto h(x_0, \tau) S(\ell + \tau),
\end{equation}
which satisfies the following:
\begin{equation}
g(\tau) = \int_0^\infty h(x_0, \tau) \mathrm{d}x_0.
\end{equation}
Finally, the cohort reproduction for this model is defined as follows:
\begin{equation}
\mathcal R_c(t) = \mathcal R_0 \int_0^\infty g(\tau) S(t+\tau) \mathrm{d} \tau.
\end{equation}

@@

First, we compare the estimates of the reproduction number $\mathcal R$ based on the intrinsic generation-interval distribution $g(\tau)$ and the forward serial-interval distribution during the exponential growth phase $f_{\textrm{\tiny exp}}(\tau)$.
Both distributions provide identical estimates of $\mathcal R$ regardless of the correlation $\rho$ between the incubation period distribution and the intrinsic generation-interval distribution (\fref{rR}A).
On the other hand, using naive serial-interval distributions that do not account for disease dynamics (i.e., assuming that the backward and the forward incubation period distributions are identical) underestimates $\mathcal R$;
as $r$ increases, ${\mathcal R}_{\textrm{\tiny naive}}$ saturates and eventually decreases due to negative serial intervals (\fref{rR}B).
While the forward serial intervals during the exponential growth phase can be also negative, the proportion of negative intervals are appropriately balanced because faster epidemic growth will lead to shorter backward incubation period (and therefore less negative serial intervals).

@@

\begin{figure}[!th]
\includegraphics[width=\textwidth]{forward.pdf}
\caption{
\textbf{Epidemiological dynamics and changes in mean forward and backward delay distributions.}
A. Cumulative incidence over time.
B-D. Changes in the mean forward incubation period, generation interval, and serial interval.
E-G. Changes in the mean backward incubation period, generation interval, and serial interval.
Black lines represent the results of a deterministic simulation.
Orange lines and points represent the average of 10 stochastic simulations;
cumulative incidence from the stochastic simulations in panel A are shown without averaging.
Forward incubation periods and intrinsic generation-intervals are assumed to be independent of each other: $h(x_0, \sigma) = k(x_0) g(\sigma)$.
See Table 1 for parameter values.
}
\label{fig:epi}
\end{figure}

\fref{epi} compares the epidemiological dynamics (A) with the mean forward (B--D) and the mean backward (E--F) delay distributions of a deterministic model based on the renewal equation (\eref{renewal}) and the corresponding stochastic realizations based on the Gillespie algorithm.
The mean forward incubation period remains constant throughout an epidemic as expected (\fref{epi}B).
The mean forward generation interval contracts as the epidemic progresses because an infected individual is less likely to infect another person as the proportion of susceptible individuals decreases (\fref{epi}C; \cite{champredon2015intrinsic}).
In contrast, the mean serial interval depends on previous incidence and therefore decreases over time (\fref{epi}D):
When incidence is increasing, symptomatic individuals are more likely to have been infected more recently (shorter backward incubation period and therefore longer forward serial interval) whereas when incidence is decreasing, symptomatic individuals are more likely to have been infected later (longer backward incubation period and therefore shorter forward serial interval).
Qualitative patterns in the changes in the mean backward delays is robust across all delay distributions because they are predominantly driven by the changes in incidence (\fref{epi}D).

\begin{figure}[!th]
\includegraphics[width=\textwidth]{forward_tease.pdf}
\caption{
\textbf{Dynamical correlations between incubation periods and the number of secondary cases.}
A. The mean forward serial interval depends on changes in the cohort reproduction number.
B. The mean backward incubation period of infectors is higher than the mean backward incubation period of all symptomatic cases.
C. Correlation between the number of secondary cases and the incubation period for individuals who got infected within the same time preiod.
D. Correlation between the number of secondary cases and the incubation period for individuals who developed symptoms within the same time preiod.
Black lines represent the results of a deterministic simulation.
Orange lines and points represent the average of 10 stochastic simulations.
Orange ribbons represent the 95\% confidence intervals.
}
\label{fig:tease}
\end{figure}

We further tease apart factors that affect the forward serial interval distribution.
\fref{tease}A compares the mean forward serial interval calculated from a deterministic model and its corresponding stochastic simulations;
failing to account for changes in $\mathcal R_c$ in \eref{forward} overestimates the mean forward serial interval.
A key distinction that needs to be made is that the forward serial-interval distributions depend on the backward incubation period of symptomatic infectors (i.e., those who were able to successfully spread the disease to another person), rather than the incubation period of all symptomatic cases (including those who were not able to spread the disease).
Therefore, the cohort size of infectors must be calculated as a product of the cohort reproduction number and incidence because cohorts with higher cohort reproduction numbers are more infectiouss and therefore more likely to be infectors.
